Aircraft Wildlife collisions

### 2021-07-14 aircraft_wildlife_doc.Rmd
Added Aircraft analysis and formatted rmd.

@@

@@

## Closer look on En Rout and Descent
```{r echo=TRUE, fig.asp=1}
#plot height density for Descent and EnRoute
ggplot(birds[birds$phase_of_flt== "Descent" | birds$phase_of_flt=="En Route",]) +
  aes(x = height, group=phase_of_flt,fill=phase_of_flt ) +
  geom_density(adjust=1.5,alpha=.8)

#Two sample t.test for height dependent on phase of flight ("Descent" and "En Route")
t.test(height~phase_of_flt, data = birds[as.integer(birds$phase_of_flt)>4 & as.integer(birds$phase_of_flt)<7 & !is.na(birds$phase_of_flt),])

@@

```{r echo=FALSE, fig.asp=1}
#relationship of amount of birds struck with clouds?
birds$effectYes<- as.factor(birds$effect != "None")
birds.filtered<-birds[!is.na(birds$effect) & !is.na(birds$sky),]

@@

```{r echo=FALSE, fig.asp=1}
#relationship of amount of birds seen with clouds?
table(birds$birds_seen)

@@

```{r echo=FALSE, fig.asp=1}
#relationship of amount of birds seen with daytime?
table(birds$birds_seen)

prop.table(table(birds.filtered$time_of_day,birds.filtered$birds_seen),1)
ggplot(birds.filtered[!is.na(birds.filtered$birds_seen) & !is.na(birds.filtered$time_of_day),])+
  aes(x=birds_seen, fill=time_of_day)+
  geom_bar(position = "fill")
```
## Birds struck out of birds seen

```{r echo=FALSE, fig.asp=1}
#Create new column
#calculate only when seen is not na
df <- birds[!is.na(birds$birds_seen) & !is.na(birds$birds_struck),]
```

check speed

```{r echo=FALSE, fig.asp=1}
# does speed has an influence on seen? the faster the less likely to see ?

spineplot(birds$speed, as.factor(is.na(birds$birds_seen)))

```
The higher the speed, the less likely to see a bird

@@

```{r echo=FALSE, fig.asp=1}
ggplot(birds.filtered[!is.na(birds.filtered$birds_struck),]) +
  aes(x = birds_struck, fill = effectYes) +
  geom_bar(position="fill")

@@

```{r echo=FALSE, fig.asp=1}
prop.table(table(birds.filtered$sky,birds.filtered$effectYes),1)
ggplot(birds[!is.na(birds$effect) & !is.na(birds$sky),])+
  aes(x=sky, fill=effectYes)+
  geom_bar(position = "fill")
```

@@

```{r echo=FALSE, fig.asp=1}
#anteil der effect=true vergleichen when seen and not seen
#...
effectYes_seen <- ggplot(birds[!is.na(birds$effectYes),]) +
  aes(x=is.na(birds_seen), fill=effectYes) +
  geom_bar(position="fill")

@@

proportion of effect is a little higher when birds were seen

dp<-ggplot(iris) +
   aes(x=Petal.Width, color=Species, fill=Species) +
   geom_density(alpha=0.25) + 
   theme(legend.position = "none") +
   scale_color_manual(values = color_pal) + 
   scale_fill_manual(values = color_pal)

bp<-ggplot(iris) +
  aes(x=Petal.Width, y=Species, fill=Species, color=Species) +
  stat_boxplot(geom="errorbar",width=0.2) +
  geom_boxplot(varwidth = TRUE, alpha=0.2) + 
  geom_jitter(alpha = 0.25, width = 0.2) +
  theme(legend.position = "none") +
  scale_color_manual(values = color_pal) + 
  scale_fill_manual(values = color_pal)

h<-ggplot(iris) +
  aes(x=Petal.Width, color=Species, fill=Species) +
  geom_histogram(bins=30, alpha=0.25) + 
  scale_color_manual(values = color_pal) + 
  scale_fill_manual(values = color_pal)

dpl<-ggplot(iris) +
   aes(x=Petal.Length, color=Species, fill=Species) +
   geom_density(alpha=0.25) + 
   theme(legend.position = "none") +
   scale_color_manual(values = color_pal) + 
   scale_fill_manual(values = color_pal)

bpl<-ggplot(iris) +
  aes(y=Petal.Length, x=Species, fill=Species, color=Species) +
  stat_boxplot(geom="errorbar",width=0.2) +
  geom_boxplot(varwidth = TRUE, alpha=0.2) + 
  theme(legend.position = "none") +
  scale_color_manual(values = color_pal) + 
  scale_fill_manual(values = color_pal)

sp<-ggplot(iris) +
  aes(x = Petal.Length, y = Petal.Width, shape = Species, color=Species, fill=Species) +
  geom_point() +
  facet_wrap(~Species) +
  scale_color_manual(values = color_pal) + 
  scale_fill_manual(values = color_pal) + 
  labs(
    title = "Petal Width and Length of 150 flowers of Iris",
    subtitle = "In relation to their species",
    x = "Petal Length in cm",
    y = "Petal Width in cm"
  ) + theme.title

dps<-ggplot(iris) +
   aes(x=Sepal.Length, color=Species, fill=Species) +
   geom_density(alpha=0.25) + 
   theme(legend.position = "none") +
   scale_color_manual(values = color_pal) + 
   scale_fill_manual(values = color_pal)

hs<-ggplot(iris) +
  aes(x=Sepal.Length, color=Species, fill=Species) +
  geom_histogram(bins=30, alpha=0.25) + 
  scale_color_manual(values = color_pal) + 
  scale_fill_manual(values = color_pal)

bps<-ggplot(iris) +
  aes(x=Sepal.Length, y=Species, fill=Species, color=Species) +
  stat_boxplot(geom="errorbar",width=0.2) +
  geom_boxplot(varwidth = TRUE, alpha=0.2) + 
  geom_jitter(alpha = 0.25, width = 0.2) +
  theme(legend.position = "none") +
  scale_color_manual(values = color_pal) + 
  scale_fill_manual(values = color_pal)

sps<-ggplot(iris) +
  aes(x = Sepal.Width, y = Sepal.Length, shape = Species, color=Species, fill=Species) +
  geom_point() +
  facet_wrap(~Species) +
  scale_color_manual(values = color_pal) + 
  scale_fill_manual(values = color_pal) + 
  labs(
    title = "Sepal Width and Length of 150 flowers of Iris",
    subtitle = "In relation to their species",
    x = "Sepal Width in cm",
    y = "Sepal Length in cm"
  ) + theme.title

(dp + h) / bp / (dpl + bpl) / sp
sps/ (dps + hs) / bps

### PCA for Iris data

To do a Principal Component Analysis we will use *prcomp()* and plot the eigenvalues in a scree plot. We see that there's an elbow at 2, so we choose 2 dimensions. *iris.pca$x* has the same dimensions as *X*.

```{r }
X<-iris[,1:4]             ## exclude the Species column
S <- cov(X)               ## compute variance-covariance matrix 
iris.pca <- prcomp(X)     ## perform PCA
lambda <- eigen(S)$values ## calculate eigenvalues
```

```{r echo=FALSE}
## scree plot
plot(lambda/sum(lambda),                        ## or instead of lambda: iris.pca$sdev^2
    main = "Scree plot of eigenvalues in %",
    col=c("black", "red", "black", "black"))
lines(lambda/sum(lambda))                       ## or: iris.pca$sdev^2
```

>>>>

sum(is.na(birds$atype)) #0
birds[,c("atype", "num_engs", "ac_mass")]
sapply(birds[,c("atype", "num_engs", "ac_mass")], function(x) sum(is.na(x)))

@@

which(birds$height==max(birds$height, na.rm = T))
birds[15053,] # >30000 foot (1 meter = 3,28084 foot)
?birds
```

## Closer look on En Rout and Descent
```{r echo=TRUE, fig.asp=0.5}
#plot height density for Descent and EnRoute
ggplot(birds[birds$phase_of_flt== "Descent" | birds$phase_of_flt=="En Route",]) +
  aes(x = height, group=phase_of_flt,fill=phase_of_flt ) +
  geom_density(adjust=1.5,alpha=.8)

@@

t.test
```{r echo=TRUE, fig.asp=1}
#Two sample t.test for height dependent on phase of flight ("Descent" and "En Route")
t.test(height~phase_of_flt, data = birds[as.integer(birds$phase_of_flt)>4 & as.integer(birds$phase_of_flt)<7 & !is.na(birds$phase_of_flt),])

@@

```{r echo=FALSE, fig.asp=0.5}
#relationship of amount of birds struck with clouds?
birds$effectYes<- as.factor(birds$effect != "None")
birds.filtered<-birds[!is.na(birds$effect) & !is.na(birds$sky),]

@@

```{r echo=FALSE, fig.asp=0.5}
#relationship of amount of birds seen with clouds?
table(birds$birds_seen)

@@

```{r echo=FALSE, fig.asp=0.5}
#relationship of amount of birds seen with daytime?
table(birds$birds_seen)

prop.table(table(birds.filtered$time_of_day,birds.filtered$birds_seen),1)
ggplot(birds.filtered[!is.na(birds.filtered$birds_seen) & !is.na(birds.filtered$time_of_day),])+
  aes(x=birds_seen, fill=time_of_day)+
  geom_bar(position = "fill")
```
check speed

```{r echo=FALSE, fig.asp=0.5}
# does speed has an influence on seen? the faster the less likely to see ?
spineplot(birds$speed, as.factor(is.na(birds$birds_seen)))
```
The higher the speed, the less likely to see a bird

### Birds struck out of birds seen

```{r echo=FALSE, fig.asp=0.5}
#Create new column
#calculate only when seen is not na
df <- birds[!is.na(birds$birds_seen) & !is.na(birds$birds_struck),]
table(birds$birds_seen, birds$birds_struck)

df$birds_struck <- factor(df$birds_struck)

#mosaic plot better than barplot visualizes also absolute differences
ggplot(df) +
  geom_mosaic(aes(x = product( birds_struck,birds_seen ), fill=birds_struck))
```

```{r echo=FALSE, fig.asp=0.5}
#Create new column
#calculate only when seen is not na
df <- birds[!is.na(birds$birds_seen) & !is.na(birds$birds_struck),]
table(birds$birds_seen, birds$birds_struck)

df$birds_struck <- factor(df$birds_struck)

levels(df$birds_seen)
levels(df$birds_struck)
df$difference_seen_strike <- as.integer(df$birds_struck) < as.integer(df$birds_seen)+2
table(df$difference_seen_strike)

barplot(prop.table(table(df$difference_seen_strike,df$effectYes),2))

ggplot(df) +
  aes(x=difference_seen_strike, fill=effectYes) +
  geom_bar()
```

@@

```{r echo=FALSE, fig.asp=0.5}
ggplot(birds.filtered[!is.na(birds.filtered$birds_struck),]) +
  aes(x = birds_struck, fill = effectYes) +
  geom_bar(position="fill")

@@

```{r echo=FALSE, fig.asp=0.5}
prop.table(table(birds.filtered$sky,birds.filtered$effectYes),1)
ggplot(birds[!is.na(birds$effect) & !is.na(birds$sky),])+
  aes(x=sky, fill=effectYes)+
  geom_bar(position = "fill")
```

@@

```{r echo=FALSE, fig.asp=0.5}
#anteil der effect=true vergleichen when seen and not seen
#...
effectYes_seen <- ggplot(birds[!is.na(birds$effectYes),]) +
  aes(x=is.na(birds_seen), fill=effectYes) +
  geom_bar(position="fill")

@@

```
proportion of effect is a little higher when birds were seen
